Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-08-25 doc/bcspaper/article.tex
Made some changes as requested by SGAI paper administrator

\title*{Partridge: An effective system for the automatic classification of the types of academic papers}
\titlerunning{Partridge: classifying academic papers}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\author{James Ravenscroft\,$^{1}$\footnote{to whom correspondence should be addressed}, Maria Liakata\,$^{2,3}$ and Amanda Clare$^{1}$}
%\address{$^{1}$Department of Computer Science, Aberystwyth University,
%$^{2}$Department of Computer Science, University of Warwick,
%$^{2}$Text mining group, EMBL-EBI Cambridge}
\author{James Ravenscroft \and Maria Liakata \and Amanda Clare}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{James Ravenscroft, Amanda Clare \at Department of Computer Science, Aberystwyth University, Aberystwyth, SY23 3DB, UK  \email{ravenscroft@papro.org.uk}, \email{afc@aber.ac.uk}
\and Maria Liakata \at Department of Computer Science, University of Warwick \email{M.Liakata@warwick.ac.uk}}

@@

\subsection*{Collection of scientific articles} 
Partridge allows users to upload any paper which is free of copyright
restrictions.  For the purpose of this study we needed a large set of papers
that we could label with CoreSC to investigate how this information assists
classification into paper type.  Open Access (OA) journals provide free read
access to their papers, but many do not permit the user of articles for data
mining purposes. The Public Library of Science (PLoS) journals contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the PubMed Central markup schema, which is compatible with
SAPIENTA, for papers published through their journal. The PLoS journals
advanced search offers approximately 50 types of paper through which to
restrict the search.  Many of these paper type categories contain too few
papers to be useful, others are too ad hoc (e.g.  {\it Message from PLoS}).
Indeed it is not clear how these paper types have been identified. We chose to
look at a range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.75\textwidth]{figures/corescs.png}
\caption{The CoreSC content of a research paper (left) and a review paper (right) randomly selected from the corpus}
\label{fig:coresc_pies} 
\end{figure}

\subsection*{Paper Processing and Classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting)

@@

A confusion matrix for the paper types is given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case Study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following F-measures:
{\bf OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 and Case Study 24.8}.

@@

\bibliographystyle{acm}
\bibliography{partridge}

>>>>

\title*{Partridge: An effective system for the automatic classification of the types of academic papers}
\titlerunning{Partridge: A system for classification of academic papers}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\author{James Ravenscroft\,$^{1}$\footnote{to whom correspondence should be addressed}, Maria Liakata\,$^{2,3}$ and Amanda Clare$^{1}$}
%\address{$^{1}$Department of Computer Science, Aberystwyth University,
%$^{2}$Department of Computer Science, University of Warwick,
%$^{2}$Text mining group, EMBL-EBI Cambridge}
\author{James Ravenscroft \and Maria Liakata \and Amanda Clare}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{James Ravenscroft, Amanda Clare \at Department of Computer Science, Aberystwyth University, Aberystwyth, SY23 3DB, UK  \email{ravenscroft@papro.org.uk}, \email{afc@aber.ac.uk}
\and Maria Liakata \at Department of Computer Science, University of Warwick \email{M.Liakata@warwick.ac.uk}}

@@

\subsection{Collection of scientific articles} 
Partridge allows users to upload any paper which is free of copyright
restrictions.  For the purpose of this study we needed a large set of papers
that we could label with CoreSC to investigate how this information assists
classification into paper type.  Open Access (OA) journals provide free read
access to their papers, but many do not permit the user of articles for data
mining purposes. The Public Library of Science (PLoS) journals contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the PubMed Central markup schema, which is compatible with
SAPIENTA, for papers published through their journal. The PLoS journals
advanced search offers approximately 50 types of paper through which to
restrict the search.  Many of these paper type categories contain too few
papers to be useful, others are too ad hoc (e.g.  {\it Message from PLoS}).
Indeed it is not clear how these paper types have been identified. We chose to
look at a range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{figures/corescs.png}
\caption{The CoreSC content of a research paper (top) and a review paper (bottom) randomly selected from the corpus}
\label{fig:coresc_pies} 
\end{figure}

\subsection{Paper Processing and Classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting)

@@

A confusion matrix for the paper types is given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case Study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following F-measures:\\
{\bf OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 and Case Study 24.8}.

@@

\bibliographystyle{spphys}
\bibliography{partridge}
